---
title: "Lab Sheet 2"
author: "STAT260, University of Otago"
date: "Sem. 2, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1 
The data that we will be using are in the file `aki.csv` on the resource page.  To download the data into R you should download the file and **place it into the directory that you are working in**.  Then in the space below call the data in R using the R code `akidata <- read.csv("aki.csv",stringsAsFactors=TRUE)`.

These data come from admissions into the ICU.  The variables measured are:

- `AKI`: presence of an acute kidney injury
- `Death_30`: did the patient die within 30 days of admission
- `uCysC`: the urine cystatin C concentration at admission 
- `sex`: 1 = female, 0 = male
- `age` (years)
- `ethnicity`
- `weight` (kg) 
- `COPD`: history of chronic pulmonary disease
- `diab`: presence of diabetes
- `hyper`: presence of hypertension

Have a look at the data using the R code `head(akidata)`.  This shows the first 6 lines of the data frame `akidata`.

_Answer:_

```{r}
akidata <- read.csv("aki.csv",stringsAsFactors=TRUE)
head(akidata)
```

******

# Task 2
Produce a histogram of the patient ages.

_Answer:_

``` {r}
hist(akidata$age)
```


******

# Task 3
Produce another histogram of the patient ages.  Ensure the x labels and plot title are well specified.

_Answer:_

```{r}
hist(akidata$age, main = "Histrogram of Patient age", xlab = "Age")
```
******

# Task 4
Produce a histogram of the patient weights.  Set the number of bins to 15.  Ensure the x labels and plot title are well specified.

_Answer:_

```{r}
patientWeightHist <- hist(akidata$weight, breaks = 15, main = "Histrogram of patient weight", xlab = "weight (kg)")
```
******

# Task 5
The y-axis is giving the frequency in each bin.  Instead of the frequency, we may prefer to have the y-axis measure the probability density (proportions).  Using `help(hist)` as a guide, produce a histogram of patient age where the y-axis is in terms of the probability density.

_Answer:_

```{r}
hist(akidata$age, freq = FALSE, main = "Histrogram of patient age (probability density)", xlab  = "Age")
```

******

# Task 6
Obtain a list of the break points and frequencies for the histogram of weight (with 15 break points).  Hint 1: use `help(hist)` as a guide.  Hint 2: you do not need to provide a plot.

```{r}
patientWeightHist$breaks
patientWeightHist$counts


```


******

# Task 7
Plot a boxplot of the patient weights.  Ensure you include a title and y-axis label.

_Answer:_

```{r}
boxplot(akidata$weight, ylab = "Weight", main = "Patient weight")
```
******

# Task 8
Plot boxplots of the patient weight by sex.  That is, the boxplot of weight for males should be beside the boxplot of weight for females.  Ensure the x and y labels are well specified.

_Answer:_

```{r}
boxplot(akidata$weight ~ akidata$sex, ylab = "Weight (kg)", xlab = "Sex (1 = Female and 0 = Male)", main = "Boxplot of patient weight by sex")
```

******
