---
title: "Lab Sheet 7"
author: "STAT260, University of Otago"
date: "Sem. 2, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


******

# Task 1 
Write three vectors.  The first contain the names of 5 friends or family members.  The second contain the heights of these 5 people.  The third contain the ages of these 5 people.  Combine these three vectors into a data frame.  Ensure that the column names are appropriately descriptive.  Note: if you don't know these values, you shoud guess them.  The object of these task is to make a data frame from the vectors, not obtaining accurate values in the vectors.


_Answer:_

```{r}
name <- c('Put', 'James', 'Julia', 'Bob', 'Jane')
height <- c(175, 165, 173, 156, 180)
age <- c(18, 21, 34, 65, 12)
family_df <- data.frame(name, height, age)
family_df
```

******

# Task 2
Using the vectors you created in task 1:

- Use R to find the age of your 2nd friend (i.e., subset the vector to get the age of the second person).  
- Use R to give you a new vector that gives the height of your 1st, 3rd and 5th friend.


_Answer:_

```{r}
family_df[2, 'age']

family_df[c(1, 3, 5), 'height']
```

******

# Task 3

Using the data frame you created in task 1:

- Use R to find the names of your friends/family over the age of 20 (feel free to change the value of 20 to something more appropriate for your data).  
- Use R to find a new data frame with only your friends/family that are taller than 1.75 metres tall (again feel free to change 1.75 to something more appropriate for your data).


_Answer:_

```{r}
family_df$name[family_df$age > 20]
family_df[family_df$height > 170, ]
```

******

# Task 4
We will use the CPS data that we have used in previous labs.  However, the data for this lab are stored in two files.

The first file is `cps_employ.csv` and contains the employment information on each subject:

- `Experience`: Number of years of work experience
- `Union`: Indicator variable for union membership (1 = member, 0 = not member)
- `Wage`: dollars per hour
- `Occupation`: 1 = management, 2 = sales, 3 = clerical, 4 = service, 5 = professional, 6 = other
- `Sector`: 0 = other, 1 = manufacturing, 2 = construction

The second file is `cps_demo.csv` and contains demographic information on each subject:

- `Education`: the number of years of education
- `South`: Indicator variable for Southern region (1 = Southern, 0 = not Southern)
- `Sex`: Indicator variable for sex (1 = female, 0 = male)
- `Age`: age (years)
- `Race`: 1 = other, 2 = hispanic, 3 = white
- `Marr`: Indicator variable for marital status (1 = married, 0 = not married)

Call each of the two data sets into R.  Combine into one data frame.  Refamiliarize yourself with the data.

_Answer:_

```{r}

cps_employ <- read.csv('./cps_employ.csv')
cps_demo <- read.csv('./cps_demo.csv')
cps_combined <- data.frame(cps_demo, cps_employ)
head(cps_combined)

```

******

# Task 5
Use `ggplot` to produce a plot comparing wages to experience for those individuals working in construction.  Hint: you should first subset out the individuals working in construction, then you should produce the appropriate plot.  Remember to label the plot appropriately.

_Answer:_

```{r}
cps_combined[cps_combined$Sector == 2, ] %>% 
  ggplot(aes(x=Experience, y=Wage))+
 geom_point()+
  geom_smooth(method = 'loess', formula = y ~ x)+
  ggtitle('Plot of wage against experience for construction workers')+
  xlab('Experience (years)')+
  ylab('Wage (USD/hour)')
```

******

# Task 6
Use `ggplot` to produce a histogram of wages for those individuals with more than 10 years of education.

_Answer:_

```{r}
cps_combined[cps_combined$Education > 10, ] %>% 
  ggplot(aes(x=Wage))+
  geom_histogram()+
  ggtitle('Histrogram of wage for individal with more than 10 years education')+
  xlab('Wage (USD/hour)')
```

******

# Task 7
Use `ggplot` to plot wages against experience for married men.  Hint: married men have both married = 1 and sex = 0.

_Answer:_

```{r}
cps_combined[cps_combined$Marriage == 1 & cps_combined$Sex == 0, ] %>% 
  ggplot(aes(x=Experience, y=Wage))+
  geom_point()+
  ggtitle('Plot of wage against expereince for married male')+
  xlab('Experience (years)')+
  ylab('Wage (USD/hour)')
```

******

# Task 8
How many hispanic women are there with wages over \$20/hour?  What about with wages under \$10/hour?

_Answer:_

```{r}
cps_combined %>% 
  mutate(hispanic_female_wage = ifelse(Wage > 20, 'Above $20/hour', ifelse(Wage < 10, 'Under $10/hour', 'Between $20 and $10/hour'))) %>% 
  filter(Race == 2 & Sex == 1) %>% 
  group_by(hispanic_female_wage) %>% 
  summarise(count = n())

```
******

# Task 9
Use `ggplot` to produce a histogram of the age of individuals who are either in management or earn over \$25/hour.  

_Answer:_

```{r}
cps_combined[cps_combined$Occupation == 1 | cps_combined$Wage > 25, ] %>% 
  ggplot(aes(x=Age)) +
  geom_histogram()+
  ggtitle('Distribution of age of individual who either in management or earn over $25/hour')+
  xlab('Age (years)')
```
