---
title: "Lab Sheet 7 - solutions"
author: "STAT260, University of Otago"
date: "Sem. 2, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


******

# Task 1 
Write three vectors.  The first contain the names of 5 friends or family members.  The second contain the heights of these 5 people.  The third contain the ages of these 5 people.  Combine these three vectors into a data frame.  Ensure that the column names are appropriately descriptive.  Note: if you don't know these values, you shoud guess them.  The object of these task is to make a data frame from the vectors, not obtaining accurate values in the vectors.


_Answer:_

```{r}
x = c("Bob","Fred", "Tom", "Dick", "Harry")
y = c(1.75, 1.88, 1.81, 1.69, 2.04)
z = c(31, 40, 28, 71, 20)
friends = data.frame(name = x, height = y, age = z)
friends
```

******

# Task 2
Using the vectors you created in task 1:

- Use R to find the age of your 2nd friend (i.e., subset the vector to get the age of the second person).  
- Use R to give you a new vector that gives the height of your 1st, 3rd and 5th friend.


_Answer:_
```{r}
z[2]
y[c(1,3,5)]
```


******

# Task 3

Using the data frame you created in task 1:

- Use R to find the names of your friends/family over the age of 20 (feel free to change the value of 20 to something more appropriate for your data).  
- Use R to find a new data frame with only your friends/family that are taller than 1.75 metres tall (again feel free to change 1.75 to something more appropriate for your data).


_Answer:_
```{r}
x[z>20] 
# OR
age20 = which(z > 20)
x[age20]

friends[friends$height>1.75,]
# OR
h175 = which(friends$height > 1.75)
friends[h175,]
```

******

# Task 4
We will use the CPS data that we have used in previous labs.  However, the data for this lab are stored in two files.

The first file is `cps_employ.csv` and contains the employment information on each subject:

- `Experience`: Number of years of work experience
- `Union`: Indicator variable for union membership (1 = member, 0 = not member)
- `Wage`: dollars per hour
- `Occupation`: 1 = management, 2 = sales, 3 = clerical, 4 = service, 5 = professional, 6 = other
- `Sector`: 0 = other, 1 = manufacturing, 2 = construction

The second file is `cps_demo.csv` and contains demographic information on each subject:

- `Education`: the number of years of education
- `South`: Indicator variable for Southern region (1 = Southern, 0 = not Southern)
- `Sex`: Indicator variable for sex (1 = female, 0 = male)
- `Age`: age (years)
- `Race`: 1 = other, 2 = hispanic, 3 = white
- `Marr`: Indicator variable for marital status (1 = married, 0 = not married)

Call each of the two data sets into R.  Combine into one data frame.  Refamiliarize yourself with the data.

_Answer:_

```{r}
cps1 = read.csv('cps_employ.csv')
cps2 = read.csv('cps_demo.csv')
str(cps1)
str(cps2)
cpsdata = data.frame(cps1,cps2)
str(cpsdata)
```

******

# Task 5
Use `ggplot` to produce a plot comparing wages to experience for those individuals working in construction.  Hint: you should first subset out the individuals working in construction, then you should produce the appropriate plot.  Remember to label the plot appropriately.

_Answer:_

```{r}
idxman = which(cpsdata$Sector == 2)
cps_man = cpsdata[idxman,]

library(ggplot2)
ggplot(data = cps_man, aes(x=Experience, y=Wage)) + geom_point() + 
  ggtitle("Construction sector") + xlab("Experience (years)") + ylab("Wage (USD/hour)")
```

******

# Task 6
Use `ggplot` to produce a histogram of wages for those individuals with more than 10 years of education.

_Answer:_

```{r}
idxed10 = which(cpsdata$Education > 10)
cps_ed10 = cpsdata[idxed10,]

ggplot(data = cps_ed10, aes(x=Wage)) + geom_histogram(bins=30) + 
  ggtitle("Individuals with more than 10 years education") + xlab("Wages (USD/hour)")
```

******

# Task 7
Use `ggplot` to plot wages against experience for married men.  Hint: married men have both married = 1 and sex = 0.

_Answer:_
```{r}
idxmm = which(cpsdata$Marriage == 1 & cpsdata$Sex == 0)
cps_man = cpsdata[idxmm,]

ggplot(data = cps_man, aes(x=Experience, y=Wage)) + geom_point() + 
  ggtitle("Married men") + xlab("Experience (years)") + ylab("Wage (USD/hour)")
```

******

# Task 8
How many hispanic women are there with wages over \$20/hour?  What about with wages under \$10/hour?

_Answer:_
```{r}
idxhw20 = which(cpsdata$Race == 2 & cpsdata$Sex == 1 & cpsdata$Wage > 20)
length(idxhw20)
idxhw10 = which(cpsdata$Race == 2 & cpsdata$Sex == 1 & cpsdata$Wage < 10)
length(idxhw10)
```

******

# Task 9
Use `ggplot` to produce a histogram of the age of individuals who are either in management or earn over \$25/hour.  

_Answer:_

```{r}
idxman25 = which(cpsdata$Occupation == 1 | cpsdata$Wage > 25)
cps_man25 = cpsdata[idxman25,]

ggplot(data = cps_man25, aes(x=Age)) + geom_histogram(bins=15) + 
  ggtitle("Individuals in management or earning more than $25/hour") + xlab("Age")
```

