---
title: "Lab Sheet 5 - solutions"
author: "STAT260, University of Otago"
date: "Sem. 2, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


******

# Task 1 
The data that we will be using are in the file `aki.csv` on the resource page (also used for lab 2).  To download the data into R you should download the file and place it into the directory that you are working in. Then in the space below, call the data in R using the R code `akidata = read.csv('aki.csv')`.

These data come from admissions into the ICU.  The variables measured are:

- `AKI`: presence of an acute kidney injury
- `Death_30`: did the patient die within 30 days of admission
- `uCysC`: the urine cystatin C concentration at admission 
- `sex`: 1 = female, 0 = male
- `age` (years)
- `ethnicity`
- `weight` (kg) 
- `COPD`: history of chronic pulmonary disease (0=none, 1=COPD)
- `diab`: presence of diabetes (0=none, 1=diabetes)
- `hyper`: presence of hypertension (0=none, 1=hypertension)

For our purposes today we will only be looking at three variables: `age`, `weight` and `sex`.  Have a look at the data using the R code `head(akidata)`.  This shows the first 6 lines of the data frame `akidata`.

_Answer:_

```{r}
library(ggplot2)
akidata = read.csv('aki.csv')
head(akidata)
```

******

# Task 2
Use `ggplot` to produce a histogram of the patient weights.

_Answer:_

```{r}
ggplot(data=akidata, aes(x=weight)) + geom_histogram()
```

******

# Task 3
Produce another histogram of the patient weights. Ensure the x labels and plot title are well specified.

_Answer:_

```{r}
ggplot(data=akidata, aes(x=weight)) + geom_histogram() + xlab("Weight (kg)") + 
  ggtitle("Weight distribution of patients admitted to ICU")
```

******

# Task 4
Produce a histogram of the patient ages.  Set the number of bins to 15 (HINT: `?geom_histogram` will load the help page), make the bars of the histogram "wheat", and the outline of the bars "brown". Ensure the x labels and plot title are well specified. Explore some different ggplot "themes" using the `+ theme_XXX()` command (where XXX is the particular theme - you can find these by typing `theme_` and waiting for RStudio to show you the various options).

_Answer:_

```{r}
ggplot(data=akidata, aes(x=age)) + 
  geom_histogram(bins=15, fill='wheat', colour='brown') + 
  theme_minimal() +
  xlab("Age (years)") + 
  ggtitle("Age distribution of patients admitted to ICU")
```

******

# Task 5
Produce a boxplot of the patient weights. Ensure the x labels and plot title are well specified. Can you find a way to make the box appear narrower in the plot?  (HINT: what are the limits of the x-axis? The `xlim` command might be useful here).

_Answer:_

```{r}
ggplot(data=akidata, aes(y=weight)) + 
  geom_boxplot() +
  ylab("Weight (kg)") + 
  ggtitle("Weight distribution of patients admitted to ICU") +
  xlim(c(-1,1))
```

******

# Task 6
Modify your plot from Task 5 to produce a boxplot of weight for each possible diabetes status (`diab` variable: 0 = "no", 1 = "yes"). Use `str` to check the data structure - since `diab` values are 0 or 1, it is considered an integer variable. Use the `as.factor()` function in your plot to convert it to a factor.

_Answer:_

```{r}
ggplot(data=akidata, aes(x=as.factor(diab), y=weight)) + 
  geom_boxplot() +
  ylab("Weight (kg)") + 
  ggtitle("Weight distribution of patients admitted to ICU")
```


