---
title: "Lab Sheet 3"
author: "STAT260, University of Otago"
date: "Sem. 2, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Task 1 
We will again be using the data in `aki.csv`.  Call this data into R in the space below.

We will be focusing on three variables in the lab today:

- `uCysC`: the urine cystatin C concentration at admission.  The interest in this study was to determine the potential for this measure to be used as a biomarker for acute kidney disease.
- `age` (years)
- `weight` (kg) 

_Answer:_

```{r}
akidata <- read.csv("../lab02/aki.csv")
```

******

# Task 2
Produce a scatterplot of the `uCysC` value (y-axis) against age (x-axis).  Ensure you label both the x and y axes (you should do this for all plots you make).

_Answer:_

```{r}
plot(akidata$age, akidata$uCysC, xlab = "Age (years)", ylab = "Biomarker (uCysC)")
```

******

# Task 3
The plot in Task 2 is not very informative because of the long right tail in the `uCysC` distribution.  Provide histograms of both `uCysC` and the log transform of `uCysC`.  Comment on the shape of these distributions.

_Answer:_

```{r}
hist(akidata$uCysC)
hist(log(akidata$uCysC))
```

The histogram on the original scale has a very long right tail.  There are a few observations that are very large.

The distribution on the log scale still does not appear symmetric or 'bell-shaped'.  It still has a right tail, but it is much shorter than that on the original scale.

******

# Task 4
Provide a scatterplot of the log of `uCysC` against `age`.  Change the colour of the points using the option `col = "goldenrod"`.  Change the point character (you can choose which character to use).

_Answer:_

```{r}
plot(akidata$age, log(akidata$uCysC), xlab = "Age (years)", ylab = "Biomarker (log uCysC)", col = "goldenrod", pch = 20)
```

******

# Task 5
Suppose that we later found another sheet of paper another 3 datapoints on it. These observations are given in `uCysC_extra` and `age_extra`:
```{r}
uCysC_extra <- c(0.06,0.30,0.13)
age_extra <- c(52,37,81)
```
Copy and paste your R code from Task 4 (to produce the scatterplot).  Then use `points` to add the extra points.  Make sure these are a different colour and point character so that you can see them.  HINT 1: The `points` function is used in the same way as the plot function with `points(x,y)`.  HINT 2: You change the colour and point character of the points in the same way as you do with the plot function. HINT 3: Remember we need all response values to be on the log scale.

_Answer:_

```{r}
plot(akidata$age, log(akidata$uCysC), xlab = "Age (years)", ylab = "Biomarker (log uCysC)", col = "goldenrod", pch = 20)
points(age_extra, log(uCysC_extra), col = "aquamarine4", pch = 17)
```

******

# Task 6
Finally, plot two additional scatterplots.  The first should be the log of `uCysC` against `weight`.  The second should be `weight` against `age`.  Do either of these plots suggest an interesting relationship between the variables?

_Answer:_


```{r}
plot(akidata$weight, log(akidata$uCysC), xlab = "Weight (kg)", ylab = "Biomarker (log uCysC)", pch = 20)
plot(akidata$age, akidata$weight, xlab = "Age (years)", ylab = "Weight (kg)", pch = 20)
```

No, neither of these plots reveals anything obvious in terms of relationships between the variables.

******
